---
title: "Question Two"
format: 
  html: 
    theme: cyborg
---

## Question Two

```{r}
#| echo: false
#| message: false
#| warning: false
library(foreach)
library(doParallel)
library(MASS)
```

```{r}
data <- galaxies
```

### Serial Bootsrap (Baseline)
This is to obtain the reference runtime

```{r}
set.seed(123)

B <- 10000
n <- length(data)

system.time({
  med_serial <- numeric(B)
  
  for (i in B) {
    sample_i <- sample(data, n, replace = TRUE)
    med_serial[i] <- median(sample_i)
    
  }
})

```

### Parallel Bootstrapping (10000 workers)


```{r}
cluster <- makeCluster(parallel :: detectCores() - 1)

# bootstrapping
set.seed(123)

system.time({
  long_par <- foreach(i = 1:B, .combine = c, .packages = "MASS") %dopar%{
    sample_i <- sample(galaxies, length(galaxies), replace = TRUE) # bootstrapping
    median(sample_i)
  }
})

# stop cluster 
stopCluster(cluster)
```

### Parallel Bootsrapping (Chunks)

Sending 10 jobs per worker

```{r}
chunks <- 10
chunk_size <- B/chunks

cluster <- makeCluster(parallel::detectCores() -1)

registerDoParallel(cluster)
  
system.time({
  par_chunk <- foreach(i = 1: chunks, .combine = c,.packages = "MASS") %dopar% {
    
    replicate(chunk_size, {
      sample_i <- sample(galaxies, length(galaxies), replace = TRUE) # bootstrapping
      median(sample_i)
    })
    
  }
})


stopCluster(cluster)
```
### Summary

The baseline bootstrap method performs really well, with a runtime that is $\approx 0$. Obviously with a larger sample size, it would take longer but I still think it is the best option of the three.

##### Between the parallel methods

The method with 10000 tasks (workers) takes quite some time in terms of efficiency. It is the least efficient method and mostly because of the overhead that comes with parallelism. There is also a matter of waiting for other tasks to finish as well which leads to some idle workers and overall long runtime.

The method with chunks compared to the one without works more efficiently, but again it still has significantly more runtime than miss sequential.
